Psychophysics of musical elements in the discrete-time representation of sound

### 2013-10-07 article/article.tex
feita ateh conclusao

There are many studies intending to model and explore a 'musical language',
or 'linguistics applied to music' or even to discern between different 'musical
idioms'~\cite{Lerdahl, Harmonia, Salzer,Alfaix}. In a simple way, a musical idiom
is the result of a chosen material together with repetition of elements and
repetition of relations between elements present along the music. The related dichotomies are outstanding, as explained at subsection~\ref{subsec:motivos}: repetition and variation, relax and tension, equilibrium and instability, consonance and dissonance, etc.

@@

The basic note was defined and characterized in quantitative terms (section~\ref{sec:notaDisc}). Next, the internal note
composition was addressed and both internal transitions and immediate treatment
were understood (section~\ref{varInternas}). Finally, this section aims at
organizing these notes in music. The gamma of resources and consequent infinitude
of possibilities is a typical situation and highly relevant for art contexts~\cite{Harmonia,Webern}.

There are studies dedicated to each one of the presented resource. For example, it is possible to obtain the 'dirty' triadic harmonies (with notes without the triad) by
superpositions of dirty fourths. Another interesting example is the simultaneous
presence of rhythms in different metrics, constituting what is
called \emph{polyrhythms}. The music piece \emph{Poli-hit mia} explores these
simultaneous metrics by impulse train (??) convolved with notes that compound the
line. Its source code is included in Appendix~\ref{ap:poli} and available online as part
of \massa.

The microtonal scales are important for the music of the 20th
century~\cite{microtonalidade} and present diverse outstanding results, like the
fourths of tone in the Indian music. The musical sequence \emph{MicroTom}
explores these resources, including microtonal melodies and microtonal harmonies
with many notes in a very reduced note scope. Its code is presented in
Appendix~\ref{ap:micro} and available online in \massa.

As pointed in subsection~\ref{subsec:mus2}, the relations between
parameters are powerful ways to acquire music pieces. The number of permuted
notes can vary during the music, revealing relationship with piece
duration. Harmonies can be made of triads (eqs.~\ref{triades}) with replicated
notes at each octaves and more numerous as minor the depth and frequency of
vibratos (eqs.~\ref{vbrGamma},~\ref{vbrAux},~\ref{vbrF},~\ref{vbrGamma2},~\ref{vbrT}),
among other uncountable possibilities.

The presented symmetries at octave divisions (eqs.~\ref{escSim}) and the
symmetries presented as permutations (Table~\ref{tab:change} and
eqs.~\ref{eq:groups}) can be used together. In the music piece \emph{3 trios}
this association is done in a systematic way in order to enable the required
listening. This is a instrumental piece, not included as a source
code~\cite{3Trios}.

The \emph{PPEPPS} (Pure Python EP: Projeto Solvente) is an EP synthesized using
the resources presented in this study. With minimal parametrization, the program
generates complete musics, allowing the easy composition of musics and sets of
musics. Using few lines of code it is possible to obtain
a directory with the generated musics. This facility and technological
demystification create aesthetically possibilities for both sharing and education.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In numerous studies, there are models and explorations of a `musical language',
many can somewhat be considered `linguistics applied to music'
and some even discern different `musical
idioms'~\cite{Lerdahl, Harmonia, Salzer,Alfaix}. In a simple way, a musical idiom
is the result of chosen material together with repetition of elements and
of relations between elements, along a music piece. In these matters, 
dichotomies are prominent, as explained in subsection~\ref{subsec:motivos}:
repetition and variation, relaxation and tension, equilibrium and instability, consonance and dissonance, etc.

@@

First, the basic note was defined and characterized in quantitative terms (section~\ref{sec:notaDisc}).
Next, internal note
composition was addressed and both internal transitions and immediate sonic treatment
were understood (section~\ref{varInternas}). Finally, this section aims at
organizing these notes in music. The gamma of resources and consequent infinitude
of praxis possibilities is a typical and highly relevant situation for artistic contexts~\cite{Harmonia,Webern}.

There are studies for each of the presented resources.
For example, it is possible to obtain `dirty' triadic harmonies (with notes out of the triad) by
superposition of perfect fourths. 
Another interesting example is the simultaneous
presence of rhythms in different metrics, constituting what is
called \emph{polyrhythm}. The music piece \emph{Poli-hit mia} explores these
simultaneous metrics by impulse trains convolved with notes.
Its source code is online as part of \massa\ toolbox~\cite{massa}.

Microtonal scales are important for 20th
century music~\cite{microtonalidade} and present diverse remarkable results, e.g.
fourths of tone in the Indian music. The musical sequence \emph{MicroTom}
explores these resources, including microtonal melodies and microtonal harmonies
with many notes in a very reduced note scope. Its code is 
available online in \massa\ toolbox~\cite{massa}.

As in subsection~\ref{subsec:mus2}, relations between
parameters are powerful ways to acquire musical pieces.
The number of permuted
notes can vary during the music, revealing relationship with piece
duration. Harmonies can be made from triads (eqs.~\ref{triades}) with replicated
notes at each octaves and more numerous as minor the depth and frequency of
vibratos (eqs.~\ref{vbrGamma},~\ref{vbrAux},~\ref{vbrF},~\ref{vbrGamma2},~\ref{vbrT}),
among other uncountable possibilities.

The symmetries at octave divisions (eqs.~\ref{escSim}) and the
symmetries presented as permutations (table~\ref{tab:change} and
eqs.~\ref{eq:groups}) can be used together. In the music piece \emph{3 trios}
this association is done in a systematic way in order to enable a dedicated
listening. This is an instrumental piece, not included as a source
code but available online~\cite{3Trios}.

\emph{PPEPPS} (Pure Python EP: Projeto Solvente) is an EP synthesized using
resources presented in this study. With minimal parametrization, the scripts
generate complete musici pieces, allowing easy composition of sets of
musics. A simple script of a few lines specifies music delivered as 16 bit
44.1kHz PCM files (WAVE). This facility and technological
demystification create aesthetical possibilities for both sharing and education.
